SunPy - Python for Solar Physics

### 2014-01-28 1-Introduction.tex

\section{Introduction}

@@

SunPy aims to provide a free, open-source and openly developed software package 
for the analysis and visualisation of solar data. SunPy makes use of the Python 
programming language and the breadth of high-quality science-focused packages 
written in that language. Python is a general-purpose, 
powerful and easy-to-learn high-level programming language.
According to the 
2014 TIOBE Index\footnote{\url{http://www.tiobe.com/index.php/content/paperinfo/tpci/index.html}},
 Python is one of the top ten most popular programming languages in the world 
and is widely used outside of scientific fields in web development, education 
and `big data' analytics.

The development of a package such as SunPy in Python is made possible by the 
rich ecosystem of scientific packages available in Python. Core packages in this 
ecosystem such as \texttt{NumPy}, \texttt{SciPy} and \texttt{matplotlib} 
provide the basic functionality expected of a scientific programming language: 
\textit{e.g.}, array manipulation, core numerical algorithms and visualisation. 
Built upon these foundations, packages such as \texttt{Astropy}, \texttt{pandas} and 
\texttt{scikit-image} provide more domain-specific functionality.

>>>>

\section{Introduction}\label{sec:Intro}

@@

SunPy aims to provide a free, open-source and openly developed software package 
for the analysis and visualisation of solar data. SunPy makes use of the Python 
programming language and the breadth of high-quality science-focused packages 
written in that language. Python is a general-purpose, 
powerful and easy-to-learn high-level programming language.
According to the 
2014 TIOBE Index\footnote{\url{http://www.tiobe.com/index.php/content/paperinfo/tpci/index.html}},
 Python is one of the top ten most popular programming languages in the world 
and is widely used outside of scientific fields in areas like web development, education 
and `big data' analytics.

The development of a package such as SunPy in Python is made possible by the 
rich ecosystem of scientific packages available in Python. Core packages in this 
ecosystem such as \texttt{NumPy}, \texttt{SciPy} and \texttt{matplotlib} 
provide the basic functionality expected of a scientific programming language -- 
\textit{e.g.}, array manipulation, core numerical algorithms and visualisation. 
Built upon these foundations, packages such as \texttt{Astropy}, \texttt{pandas} and 
\texttt{scikit-image} provide more domain-specific functionality.

### 2014-01-28 5-Dev.tex

\section{Development and Community}
SunPy is a community-developed library, designed and developed for and by 
the solar physics community. Not only is all the source code publicly available 
online under the permissive 2-clause BSD licence, the whole 
development 
process is also online and open for anyone to contribute to.
SunPy's development makes use of the online 
service GitHub (\url{http://github.com}) and Git\footnote{\url{http://git-scm.com/}}
as its distributed version control software.

@@

Many of the larger features in SunPy have been developed 
with the support of both the ESA's Summer of Code in Space (SOCIS) and the Google
Summer of Code (GSOC). Initial development of SunPy's VSO and HEK 
implementations were funded by the SOCIS program as well as a prototyped GUI 
and N-Dimensional datatype implementations. SunPy was supported by GSOC in 2013 
through the Python Software Foundation (PSF). It supported development of the HELIO and 
database sub-packages.

\subsection{Development Workflow}
The development workflow that GitHub enables is simple and transparent. The 
primary source code repository \url{http://github.com/sunpy/sunpy} can be 
`forked' by anyone with a GitHub account who can then modify their fork as they 
wish. Upon completion of any modifications it is possible for a user to create 
a 'Pull Request' (\url{http://github.com/sunpy/sunpy/pulls}) to the main SunPy 
repository in order to incorporate their modification into the Sunpy code base. 
Pull requests are reviewed by the SunPy community before being accepted. Community consensus
is required to add new features.

>>>>

\section{Development and Community}\label{sec:dev}
SunPy is a community-developed library, designed and developed for and by 
the solar physics community. Not only is all the source code publicly available 
online under the permissive 2-clause BSD licence, the whole 
development 
process is also online and open for anyone to contribute to.
SunPy's development makes use of the online 
service GitHub (\url{http://github.com}) and Git\footnote{\url{http://git-scm.com/}}
as its distributed version control software.

@@

Many of the larger features in SunPy have been developed 
with the support of both the ESA-SOCIS and GSOC. 
Initial development of SunPy's VSO and HEK 
implementations were funded by the SOCIS program as well as a prototyped GUI 
and N-Dimensional datatype implementations. SunPy was supported by GSOC in 2013 
through the Python Software Foundation (PSF). It supported development of the HELIO and 
database sub-packages.

\subsection{Development Workflow}
The development workflow that GitHub enables is simple and transparent. The 
primary source code repository \url{http://github.com/sunpy/sunpy} can be 
`forked' by anyone with a GitHub account who can then modify their fork as they 
wish. Upon completion of any modifications it is possible for a user to create 
a 'Pull Request' (\url{http://github.com/sunpy/sunpy/pulls}) to the main SunPy 
repository in order to incorporate their modification into the SunPy code base. 
Pull requests are reviewed by the SunPy community before being accepted. Community consensus
is required to add new features.

### 2014-01-28 6-Future.tex

\section{Future of SunPy}
SunPy as a project has existed for roughly three years. In this time the code 
base has grown to over $110,000$ lines. SunPy in it's current form is a useful 
package for the analysis of calibrated solar data, however, the code base is 
still very changeable from release to release.

The near-term future of the SunPy library is likely to continue this changeable 
nature with re-factoring and restructuring planned, especially within the 
core data types. The main focus upon the data types defined inside sunpy is to 
improve the interoperability and to redesign them in a more consistent manner.

In the past 12 months the collaboration with the Astropy project 
\cite{theastropycollaboration2013} has grown. The future of SunPy 
will continue this collaboration as the Astropy package provides many core 
tools that are applicable to all types of astronomy. For example the 
\texttt{astropy.units} submodule provides a simple and efficient way of 
representing physical units in Python data types, there are plans to 
incorporate this widely in the SunPy code-base.

>>>>

\section{Future of SunPy}\label{sec:future}
SunPy as a project has existed for roughly three years. In this time 
the code 
base has grown to over 110,000 lines. SunPy in its current form is a 
useful 
package for the analysis of calibrated solar data, however, the code 
base is 
still very changeable from release to release.

As discussed in Section \ref{sec:Intro}, the primary focus of the 
\texttt{sunpy} library is the analysis and visualisation of `high-level' solar 
data. This means data that has been put through instrument processing 
and 
calibration routines, and contains full (WCS) coordinate information. 
The 
development plan for SunPy is to continue development within this 
scope. The 
primary components of this plan are to provide a set of data types 
that are 
interchangeable with one another: \textit{e.g.}, if you slice a 
\texttt{MapCube} 
along one coordinate, a \texttt{LightCurve} of intensity along the 
time range of 
the \texttt{MapCube} should be returned. To achieve this goal, all the 
data 
types need to share a unified coordinate system architecture so that 
each data 
type is aware of what the physical type of its data is and how 
operations on 
that data should be performed.

In concert with the work on the data types, further integration with 
the 
\texttt{astropy} package will enable SunPy to inherit the large 
amount of work 
done in that project to support a platform for astronomical data 
analysis. For 
example, the \texttt{astropy.units} submodule provides a simple and 
efficient 
way of representing physical units in Python data types; there are 
plans to 
incorporate this widely in the SunPy code base. In the past 12 months, 
the 
collaboration with the Astropy project 
\citep{theastropycollaboration2013} and 
its community has grown, abd this direct collaboration between Astropy 
% what abd means?
and 
SunPy will be one of the key strengths of SunPy going forward.
